---
title: "FPP3"
author: "E9"
date: "4/17/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Q1

## a
They all indicate that the data are white noise, as we expect 95% of the spikes in the ACF to lie within ±2/√T ± 2 / T where T is the length of the time series.

## b

Ts are different.


# Q2

```{r echo=FALSE, message=FALSE, warning=FALSE, Load_packages}

library(fpp3)


```

2. A classic example of a non-stationary series is the daily closing IBM stock price series (data set ibmclose). Use R to plot the daily closing prices for IBM stock and the ACF and PACF. Explain how each plot shows that the series is non-stationary and should be differenced.

```{r echo=FALSE, message=FALSE, warning=FALSE, Question2}

google_2015 <- gafa_stock %>%
  filter(Symbol == "AMZN") 
google_2015  %>% autoplot()
google_2015 %>% ACF(Close) %>% 
  autoplot() + labs(subtitle = "AMZN closing stock price")
google_2015 %>% ACF(difference(Close)) %>% 
  autoplot() + labs(subtitle = "Changes in AMZN cl")

```


3. 


```{r}
global_economy %>% filter(Code == 'TUR') -> df
df %>%
  features(GDP %>% log() %>% difference(), features = unitroot_kpss)
```




```{r}
aus_accommodation %>% filter(State  == 'Tasmania') -> df
df %>%
  features(Takings %>% log() %>% difference(), features = unitroot_kpss)
```



```{r}
souvenirs  -> df
df %>%
  features(Sales %>% log() %>% difference(), features = unitroot_kpss)
```





### first equation : wt = (yt^(0.5) - 1)/(0.5)
### second equation : (1 - B)(1 - B^12)wt = et, where et is a white noise series.


5. For your retail data (from Exercise 3 in Section 2.10), find the appropriate order of differencing (after transformation if necessary) to obtain stationary data.

```{r echo=FALSE, message=FALSE, warning=FALSE, Question5}

set.seed(12345678)
myseries <- aus_retail %>%
  filter(`Series ID` == sample(aus_retail$`Series ID`,1))

library(forecast)
BoxCox.lambda(myseries$Turnover) ->lambda


myseries %>%
  features(box_cox(Turnover, lambda) %>%
             difference(), features = unitroot_kpss)


myseries %>% autoplot((box_cox(Turnover, lambda) %>%
             difference()))

```

6. Use R to simulate and plot some data from simple ARIMA models.



## a


```{r}
y <- numeric(100)
e <- rnorm(100)
for(i in 2:100)
  y[i] <- 0.6*y[i-1] + e[i]
sim <- tsibble(idx = seq_len(100), y = y, index = idx)
```

## b


```{r}
autoplot(sim)
```



larger phi leads to higher variation.

